---
title: "Reproducible Research"
author: ""
date: ""
output: html_document
---

```{r}
#Loading and pre-processing the data
data <-read.csv("C:/Users/vvkal/OneDrive/Documents/MBA/DataScience-Capstone/Reproducible_Research_data_activity/activity.csv")
```

#What is mean total number of steps taken per day?
#For this part of the assignment, you can ignore the missing values in the dataset.
#Calculate and report the mean and median total number of steps taken per day
```{r}
stepsByDay <- tapply(data$steps, data$date, sum, na.rm=TRUE)
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```

#Make a histogram of the total number of steps taken each day

```{r}
library("dplyr")
library("ggplot2")
```

```{r}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

```{r}
#What is the average daily activity pattern?
avgStepsPerTimeBlock <- aggregate(x=list(meanSteps=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
#Time- series
ggplot(data=avgStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```

```{r}
#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
mostSteps <- which.max(avgStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", avgStepsPerTimeBlock[mostSteps,'interval'])
```


```{r}
#Imputing missing values
#Calculate and report the total number of missing values in the dataset
numMissingValues <- length(which(is.na(data$steps)))
print(numMissingValues)
#Number of missing values: 2304
```

#2. Devise a strategy for filling in all of the missing values in the dataset.
#Steps column has NA values that need to be imputed
#Missing values can be imputed either with mean, median, most_frequent or constant data
#Here we will use mean
#3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
dataImputed <- data
dataImputed$steps[is.na(dataImputed$steps)] <- mean(dataImputed$steps, na.rm = TRUE)
```

```{r}
#4. Make a histogram of the total number of steps taken each day
stepsByDayImputed <- tapply(dataImputed$steps, dataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

#Are there differences in activity patterns between weekdays and weekends?
```{r}
dataImputed$dateType <-  ifelse(as.POSIXlt(dataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

#Make a panel plot containing a time series plot

```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=dataImputed, mean)

ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")

```
